---
title: "p8105_hw3_ms5965"
output: github_document
---

#### Initial Set-up

```{r setup}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1 

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in orders by user. There are user/order variables -- user ID, order ID, order day and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 


#### How many aisles, and which, are the most items ordered from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

#### Making a plot 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust = 1))
```

#### Making a table

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  
knitr::kable()
```

#### Table for Pink Lady Apples and Coffee Ice Cream

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour = mean(order_hour_of_day)) %>% 
  
  mutate(order_dow = recode(order_dow,
                "0" = "Sunday", 
                "1" = "Monday",
                "2" = "Tuesday",
                "3" = "Wednesday",
                "4" = "Thursday",
                "5" = "Friday",
                "6" = "Saturday")) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

## Problem 2

### Part A

Importing, tidying, and describing the dataset. 

```{r}
accel_data = read_csv(file = "./data/accel_data.csv") %>% 
janitor::clean_names() 

tidy_accel_data = accel_data %>% 
  
pivot_longer(
  activity_1:activity_1440,
  names_to = "minutes",
  names_prefix = "activity_",
  values_to = "activity_count") %>% 
  
  drop_na(activity_count) %>% 
  
mutate(
  dow = recode(day,
               "Monday" = "weekday",
               "Tuesday" = "weekday",
               "Wednesday" = "weekday",
               "Thursday" = "weekday",
               "Friday" = "weekday",
               "Saturday" = "weekend",
               "Sunday" = "weekend"),
 
   day = factor(day, levels = c(
     "Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")),
  activity_count = as.numeric(activity_count),
  minutes = as.numeric(minutes))
```

This dataset has `r nrow(tidy_accel_data)` rows and `r ncol(tidy_accel_data)` columns. 

The dataset contains the following information, collected from an _accelerometer_, about the 65 year old male patient admitted at the Advanced Cardiac Care Center of Columbia University Medical Center:

*   Week number, day of the week and the day number since the start of the data collection. 

*   The observations under the activity_count variable are the number of activities per minute. 

*   The minutes variable represents each minute of a 24-hour day starting at midnight. 

### Part B

```{r}
tidy_accel_data %>% 
  group_by(week, day) %>% 
  mutate(total_per_day = sum(activity_count)) %>% 
  knitr::kable(digits = 0)  
```

The least amount of activity was in week four. The patient consistently does a higher amount of activity on Fridays and on the last two Saturdays, he had the least amount of activity. 

### Part C

```{r}
tidy_accel_data %>%
  ggplot(aes(x = minutes, y = activity_count)) +
  geom_line(aes(color = day), alpha = 0.5)+
  geom_smooth(aes(color = day), alpha = .5) + 
  scale_x_continuous(
    breaks = c(0, 240, 480, 720, 960, 1200, 1440)) +
labs(
  title = "",
  x = "Minutes",
  y = "Activity Count"
) + 
  theme(legend.position = "bottom") 
```

There is low activity on either extremes of the day - early morning or late at night. Majority of the activity is between 1000 and 1300 minutes. 
